---
title: "Task02"
format: html
editor: visual
---

Task 2: Activation Function Analysis

---

## **2.1 Visualizing `swish()` Activation Function and its Gradient**

```{r}
#| echo: true
library(tidyverse)
```

You can add options to executable code like this

```{r}

#| eval: true
#| echo: true
#| label: fig-swish
#| fig-cap: "Swish Activation Function"
#| fig-width: 7
#| fig-height: 4

swish <- function(x) {
  x / (1 + exp(-x))
}

x <- -10:10
y <- swish(x)
tbl_swish <- tibble(x, y)

ggplot(tbl_swish, aes(x, y)) +
  geom_line(color = "darkblue", linewidth = 1) +
  labs(x = "x", y = "swish(x)")

```

The `echo: false` option disables the printing of code (only output is displayed).

```{r}
#| eval: true
#| echo: true
#| output: true
#| label: fig-d-swish
#| fig-cap: "Gradient of Swish Activation Function"
#| fig-width: 7
#| fig-height: 4

d_swish <- function(x) {
  sigmoid <- 1 / (1 + exp(-x))
  sigmoid + x * sigmoid * (1 - sigmoid)
}

x <- -10:10
y <- d_swish(x)
tbl_d_swish <- tibble(x, y)

ggplot(tbl_d_swish, aes(x, y)) +
  geom_line(color = "darkblue", linewidth = 1) +
  labs(x = "x", y = "d_swish(x)")

```

Unlike ReLU, whose gradient is strictly 0 for negative inputs and 1 for positive inputs, the gradient of Swish is smooth and non-zero for most inputs, with values that gradually transition rather than being abruptly cut off. This helps avoid the dying neuron problem seen in ReLU, where neurons with negative inputs stop learning entirely. However, Swish is computationally more expensive as it requires calculating the sigmoid function, while ReLU involves only a simple threshold operation.

## **2.2 Toy Examples Calling `swish()` and `relu()`**

```{r}
#| eval: true
#| echo: true

student_ids <- c(2501174, 2501901, 2500511, 2503678, 2501103)

# Replace with the median of your team's student IDs
set.seed(2500511)  # <-- change this to your team's median student ID

# Define relu() if not already defined
relu <- function(x) {
  ifelse(x > 0, x, 0)
}

# Generate 100 random values between -10 and 10
x <- runif(100, min = -10, max = 10)

# Create tibble
tbl_activations <- tibble(
  x        = x,
  swish_x  = round(swish(x), 2),
  relu_x   = round(relu(x), 2)
)

# Display first 10 rows
head(tbl_activations, 10)
```

```{r}
#| eval: true
#| echo: true

# Add is_dead column
tbl_activations <- tbl_activations |>
  mutate(is_dead = relu_x == 0)

# Display first 10 rows with selected columns
tbl_activations |>
  select(x, swish_x, relu_x, is_dead) |>
  head(10)

# Summary table
tbl_summary <- tibble(
  relu_dead_count = sum(tbl_activations$is_dead),
  swish_mean      = round(mean(tbl_activations$swish_x), 2),
  swish_sd        = round(sd(tbl_activations$swish_x), 2),
  relu_mean       = round(mean(tbl_activations$relu_x), 2),
  relu_sd         = round(sd(tbl_activations$relu_x), 2)
)

tbl_summary
```
The statistics support the claim that Swish avoids dead neurons. While ReLU produces exactly zero for all negative inputs while  being reflected by a non-zero relu_dead_count of approximately 50 out of 100 samples, Swish returns small but non-zero values for negative inputs. Swish also shows a higher mean and standard deviation compared to ReLU, indicating more active and varied neuron outputs, which promotes better gradient flow and more effective learning during training.

```{r}
#| eval: true
#| echo: true

# Add is_dead column
tbl_activations <- tbl_activations |>
  mutate(is_dead = relu_x == 0)

# Display first 10 rows with selected columns
tbl_activations |>
  select(x, swish_x, relu_x, is_dead) |>
  head(10)

# One-row summary table
tbl_summary <- tbl_activations |>
  summarise(
    relu_dead_count = sum(is_dead),
    swish_mean      = round(mean(swish_x), 2),
    swish_sd        = round(sd(swish_x), 2),
    relu_mean       = round(mean(relu_x), 2),
    relu_sd         = round(sd(relu_x), 2)
  )

tbl_summary
```

```{r}
#| eval: true
#| echo: true

# Reshape to long format
tbl_long <- tbl_activations |>
  select(x, swish_x, relu_x) |>
  pivot_longer(
    cols      = c(swish_x, relu_x),
    names_to  = "function_name",
    values_to = "activation_value"
  )

# Display first 10 rows
head(tbl_long, 10)

# Summary table grouped by function_name
tbl_long |>
  group_by(function_name) |>
  summarise(
    mean = round(mean(activation_value), 2),
    sd   = round(sd(activation_value), 2),
    min  = round(min(activation_value), 2),
    max  = round(max(activation_value), 2)
  )

```
