---
title: "task2"
format: html
---

---

---

```{r}
# Load libraries
library(tidyverse)
library(ggplot2)
library(scales)
library(readxl)
library(tidyquant)
```

```{r}


monthly_dj <- read_excel("dja-performance-report-monthly.xlsx", skip = 5,          # skips rows 1-5 (blank rows + "As of:" metadata)
  col_names = TRUE
  )

monthly_dj <- monthly_dj %>%
  mutate(
    date = as.Date(monthly_dj$"Effective Date", format = "%m/%d/%Y"),
    close_value = as.numeric(monthly_dj$"Close Value")
  ) %>%
  select(-`Effective Date`, -`Close Value`) %>%
  arrange(date)
  


head(monthly_dj)
class(monthly_dj)
```

```{r}
gap_check <- monthly_dj %>%
  arrange(date) %>%
  mutate(
    month = as.integer(format(date, "%m")),
    year  = as.integer(format(date, "%Y")),
    months_to_next_row = (lead(year) - year) * 12 + (lead(month) - month)
  )

# Rows where the *next* observation is more than 1 month away
# — this row is the gap's starting point
gaps <- gap_check %>%
  filter(months_to_next_row > 1)%>%
  select(`date`, `months_to_next_row`)

gaps
```

Reason:

The NYSE closed July 31, 1914–December 12, 1914, its longest shutdown ever, after WWI's outbreak triggered fears of a foreign sell-off and gold drain from the US, forcing officials to halt trading rather than let panic selling collapse the market.

```{r}
monthly_dj <- monthly_dj %>%
  arrange(date) %>%
  mutate(
    rel_change_pct = (close_value - lag(close_value)) / lag(close_value) * 100,
    month_abb = factor(month.abb[as.integer(format(date, "%m"))],
                        levels = month.abb, ordered = TRUE)
  )

```

```{r}
medians <- monthly_dj %>%
  filter(!is.na(rel_change_pct)) %>%
  group_by(month_abb) %>%
  summarise(median_val = median(rel_change_pct), .groups = "drop")

ggplot(monthly_dj %>% filter(!is.na(rel_change_pct)),
       aes(x = month_abb, y = rel_change_pct)) +
  geom_violin(fill = "grey85", color = "grey40", trim = TRUE) +
  geom_boxplot(width = 0.15, outlier.shape = NA, fill = "white") +
  geom_line(data = medians, aes(x = month_abb, y = median_val, group = 1),
            color = "firebrick", linewidth = 0.8) +
  geom_point(data = medians, aes(x = month_abb, y = median_val),
             color = "firebrick", size = 1.5) +
  labs(title = "DJIA Monthly Relative Change (%)",
       x = "Month", y = "Relative Change (%)") +
  theme_minimal()

# --- Visualization B: faceted freqpoly + rug + mean vline ---
means <- monthly_dj %>%
  filter(!is.na(rel_change_pct)) %>%
  group_by(month_abb) %>%
  summarise(mean_val = mean(rel_change_pct), .groups = "drop")

ggplot(monthly_dj %>% filter(!is.na(rel_change_pct)),
       aes(x = rel_change_pct)) +
  geom_freqpoly(bins = 10, color = "steelblue") +
  geom_rug(sides = "b", color = "steelblue4") +
  geom_vline(data = means, aes(xintercept = mean_val),
             linetype = "dashed", color = "firebrick") +
  facet_wrap(~ month_abb, nrow = 4, ncol = 3) +
  labs(title = "DJIA Monthly Relative Change (%) by Month",
       x = "Relative Change (%)", y = "Count") +
  theme_minimal()
```

```{r}
ticker_btc <- "BTC-USD"
ticker_gold <- "GC=F"
start_date <- "2016-01-01"
end_date <- "2026-06-30"

# Convert to monthly data
monthly_btc <- tq_get(ticker_btc, from = start_date, to = end_date) |>
  drop_na(adjusted) |>
  tq_transmute(select = adjusted, mutate_fun = to.period, period = "months")

monthly_gold <- tq_get(ticker_gold, from = start_date, to = end_date) |>
  drop_na(adjusted) |>
  tq_transmute(select = adjusted, mutate_fun = to.period, period = "months")

head(monthly_btc, 10)
head(monthly_gold, 10)
```

```{r}

combined <- bind_rows(
  monthly_dj   %>% transmute(month = floor_date(date, "month"), close_value, asset = "DJIA"),
  monthly_btc  %>% transmute(month = floor_date(date, "month"), close_value = adjusted, asset = "BTC"),
  monthly_gold %>% transmute(month = floor_date(date, "month"), close_value = adjusted, asset = "Gold")
) %>%
  arrange(asset, month)

head
```

```{r}

ggplot(combined, aes(x = month, y = close_value, color = asset)) +
  geom_line(linewidth = 0.9) +
  scale_y_log10(labels = label_dollar()) +          # log scale keeps all 3 curves visible/comparable
  scale_x_date(date_breaks = "5 year", date_labels = "%Y") +
  scale_color_manual(values = c("BTC" = "#F7931A",   # optional: brand-ish colors
                                 "ETH" = "#627EEA",
                                 "XRP" = "#23292F")) +
  labs(
    title = "Monthly Closing Prices of BTC, ETH, and XRP",
    subtitle = "Log scale used to make trends comparable across very different price ranges",
    x = "Month",
    y = "Closing Price (USD, log scale)",
    color = "Asset"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

```{r}
## ------------------------------------------------------------------------
## 1. compute_drawdown()
##    Peak-to-trough decline (%) at every point in time, using the running
##    (all-time-so-far) maximum as the "peak" reference at each step.
## ------------------------------------------------------------------------
compute_drawdown <- function(prices) {
  running_max <- cummax(prices)                    # highest price seen up to t
  drawdown <- (running_max - prices) / running_max * 100
  return(drawdown)                                  # vector: drawdown % at each t
}

## ------------------------------------------------------------------------
## 2. Verify on the test vector (expected largest drawdown = 40%)
## ------------------------------------------------------------------------
prices_test <- c(100, 60, 120, 90)
compute_drawdown(prices_test)
# -> 0  40  0  25   =>  max(.) = 40   ✓ matches the expected 40%

## ------------------------------------------------------------------------
## 3. Load the real price series (monthly close, past 10 years)
## ------------------------------------------------------------------------

# --- DJIA (from dja-performance-report-monthly.xlsx) ---
dj_raw <- read_excel("dja-performance-report-monthly.xlsx",
                      sheet = "dja-performance-report-monthly", skip = 5) %>%
  filter(`Index Name` == "Dow Jones Industrial Average") %>%
  mutate(`Effective Date` = as.Date(`Effective Date`, format = "%m/%d/%Y")) %>%
  arrange(`Effective Date`)

decade_start <- max(dj_raw$`Effective Date`) %m-% years(10)
dj <- dj_raw %>% filter(`Effective Date` >= decade_start)

# --- Gold (monthly close, same 10-year window) ---
# TODO: point this at your gold price file, e.g.:
# gold <- read_excel("gold-performance-report-monthly.xlsx", skip = 5) %>% ...
gold <- NULL

# --- BTC (monthly close, same 10-year window) ---
# TODO: point this at your BTC price file, e.g.:
# btc <- read_excel("btc-performance-report-monthly.xlsx", skip = 5) %>% ...
btc <- NULL

## ------------------------------------------------------------------------
## 4. Apply compute_drawdown() to each series
## ------------------------------------------------------------------------
drawdowns_dj   <- compute_drawdown(dj$`Close Value`)
# drawdowns_gold <- compute_drawdown(gold$Close)   # uncomment once gold is loaded
# drawdowns_btc  <- compute_drawdown(btc$Close)    # uncomment once btc is loaded

biggest_drawdown_dj   <- max(drawdowns_dj)
# biggest_drawdown_gold <- max(drawdowns_gold)
# biggest_drawdown_btc  <- max(drawdowns_btc)

biggest_drawdown_dj
# biggest_drawdown_gold
# biggest_drawdown_btc

# Peak / trough dates for the DJIA's largest drawdown
trough_i <- which.max(drawdowns_dj)
peak_i   <- max(which(dj$`Close Value`[1:trough_i] == cummax(dj$`Close Value`)[trough_i]))
dj$`Effective Date`[peak_i]    # peak month
dj$`Effective Date`[trough_i]  # trough month

```

For DJIA I have a firm, verified answer from your actual data. For Gold and BTC I still don't have your script's output — and I want to flag why that matters here rather than guess: public sources disagree on exact monthly figures (BTC's 2017–18 crash and 2021–22 crash are both cited as its "largest," in the 73–84% range depending on methodology; gold's 2022 correction is cited anywhere from 20–23%), so pulling a number from search risks being wrong for your specific adjusted-price, monthly-close series.\
What I can confirm now — DJIA:\
\
Largest drawdown: 23.2%\
Peak: December 2019 → Trough: March 2020\
Event: the COVID-19 crash — markets peaked just before the pandemic was declared, then fell sharply as lockdowns began.
