---
title: "Task02"
format: html
editor: visual
---

Visualizing swish() Activation Function and its Gradient Toy Examples Calling swish() and relu() Exploring Other Activation Functions ---


## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
#| echo: true
library(tidyverse)
```

You can add options to executable code like this

```{r}

#| eval: true
#| echo: true
#| label: fig-swish
#| fig-cap: "Swish Activation Function"
#| fig-width: 7
#| fig-height: 4

swish <- function(x) {
  x / (1 + exp(-x))
}

x <- -10:10
y <- swish(x)
tbl_swish <- tibble(x, y)

ggplot(tbl_swish, aes(x, y)) +
  geom_line(color = "darkblue", linewidth = 1) +
  labs(x = "x", y = "swish(x)")

```

The `echo: false` option disables the printing of code (only output is displayed).

```{r}
#| eval: true
#| echo: true
#| output: true
#| label: fig-d-swish
#| fig-cap: "Gradient of Swish Activation Function"
#| fig-width: 7
#| fig-height: 4

d_swish <- function(x) {
  sigmoid <- 1 / (1 + exp(-x))
  sigmoid + x * sigmoid * (1 - sigmoid)
}

x <- -10:10
y <- d_swish(x)
tbl_d_swish <- tibble(x, y)

ggplot(tbl_d_swish, aes(x, y)) +
  geom_line(color = "darkblue", linewidth = 1) +
  labs(x = "x", y = "d_swish(x)")

```








