---
title: "Task02"
format: html
editor: visual
---

Task 2: Activation Function Analysis

------------------------------------------------------------------------

## **2.1 Visualizing `swish()` Activation Function and its Gradient**

```{r}
#| echo: true
library(tidyverse)
```

You can add options to executable code like this

```{r}

#| eval: true
#| echo: true
#| label: fig-swish
#| fig-cap: "Swish Activation Function"
#| fig-width: 7
#| fig-height: 4

swish <- function(x) {
  x / (1 + exp(-x))
}

x <- -10:10
y <- swish(x)
tbl_swish <- tibble(x, y)

ggplot(tbl_swish, aes(x, y)) +
  geom_line(color = "darkblue", linewidth = 1) +
  labs(x = "x", y = "swish(x)")

```

The `echo: false` option disables the printing of code (only output is displayed).

```{r}
#| eval: true
#| echo: true
#| output: true
#| label: fig-d-swish
#| fig-cap: "Gradient of Swish Activation Function"
#| fig-width: 7
#| fig-height: 4

d_swish <- function(x) {
  sigmoid <- 1 / (1 + exp(-x))
  sigmoid + x * sigmoid * (1 - sigmoid)
}

x <- -10:10
y <- d_swish(x)
tbl_d_swish <- tibble(x, y)

ggplot(tbl_d_swish, aes(x, y)) +
  geom_line(color = "darkblue", linewidth = 1) +
  labs(x = "x", y = "d_swish(x)")

```

Unlike ReLU, whose gradient is strictly 0 for negative inputs and 1 for positive inputs, the gradient of Swish is smooth and non-zero for most inputs, with values that gradually transition rather than being abruptly cut off. This helps avoid the dying neuron problem seen in ReLU, where neurons with negative inputs stop learning entirely. However, Swish is computationally more expensive as it requires calculating the sigmoid function, while ReLU involves only a simple threshold operation.

## **2.2 Toy Examples Calling `swish()` and `relu()`**

```{r}
#| eval: true
#| echo: true

student_ids <- c(2501174, 2501901, 2500511, 2503678, 2501103)

student_ids <- c(2501174, 2501901, 2500511, 2503678, 2501103)
seed <- as.integer(median(student_ids))
# Replace with the median of your team's student IDs
set.seed(seed)  # <-- change this to your team's median student ID

# Define relu() if not already defined
relu <- function(x) {
  ifelse(x > 0, x, 0)
}

# Generate 100 random values between -10 and 10
x <- runif(100, min = -10, max = 10)

# Create tibble
tbl_activations <- tibble(
  x        = x,
  swish_x  = round(swish(x), 2),
  relu_x   = round(relu(x), 2)
)

# Display first 10 rows
head(tbl_activations, 10)
```

```{r}
#| eval: true
#| echo: true

# Add is_dead column
tbl_activations <- tbl_activations |>
  mutate(is_dead = relu_x == 0)

# Display first 10 rows with selected columns
tbl_activations |>
  select(x, swish_x, relu_x, is_dead) |>
  head(10)

# Summary table
tbl_summary <- tibble(
  relu_dead_count = sum(tbl_activations$is_dead),
  swish_mean      = round(mean(tbl_activations$swish_x), 2),
  swish_sd        = round(sd(tbl_activations$swish_x), 2),
  relu_mean       = round(mean(tbl_activations$relu_x), 2),
  relu_sd         = round(sd(tbl_activations$relu_x), 2)
)

tbl_summary
```

The statistics support the claim that Swish avoids dead neurons. While ReLU produces exactly zero for all negative inputs while being reflected by a non-zero relu_dead_count of approximately 50 out of 100 samples, Swish returns small but non-zero values for negative inputs. Swish also shows a higher mean and standard deviation compared to ReLU, indicating more active and varied neuron outputs, which promotes better gradient flow and more effective learning during training.

```{r}
#| eval: true
#| echo: true

# Reshape to long format
tbl_long <- tbl_activations |>
  select(x, swish_x, relu_x) |>
  pivot_longer(
    cols      = c(swish_x, relu_x),
    names_to  = "function_name",
    values_to = "activation_value"
  )

# Display first 10 rows
head(tbl_long, 10)

# Summary table grouped by function_name
tbl_long |>
  group_by(function_name) |>
  summarise(
    mean = round(mean(activation_value), 2),
    sd   = round(sd(activation_value), 2),
    min  = round(min(activation_value), 2),
    max  = round(max(activation_value), 2)
  )

```

## **2.3 Exploring Activation Functions**

```{r}
#| eval: true
#| echo: true

# Define ELU function (alpha = 1 by default)
elu <- function(x, alpha = 1) {
  ifelse(x > 0, x, alpha * (exp(x) - 1))
}

# Add elu_x column to existing tibble
tbl_activations <- tbl_activations |>
  mutate(elu_x = round(elu(x), 2))

# Display first 10 rows with selected columns
tbl_activations |>
  select(x, swish_x, relu_x, elu_x) |>
  head(10)
```

Description of other activation functions:
ELU is an activation function that outputs the input directly if positive, but uses an exponential curve for negative values to smooth gradients and prevent inactive neurons


## 3. Reflection on the Use of Generative AI

##** 3.1 Gen AI Tool used & Rationale**
Our team used Claude 4.6 Sonnet for this assessment. We chose it because it demonstrates better performance in handling programming languages and code logic compared to other available AI tools.

##** 3.2 Examples of usage of GenAI tool**
We utilized Claude as a quality-assurance and documentation partner:

Code Verification: We used it to check our syntax and verify code outputs, catching edge cases early.

Documentation: It helped write clear code comments and descriptions, translating complex logic into easily readable documentation.

##** 3.3 Key Benefits**
The tool acted as an effective teaching aid. Instead of just fixing code, it explained core programming concepts by pinpointing specific code lines. This deepened our understanding of our own codebase.

##  3.4 Challenges & Limitations
The main challenge was managing AI hallucinations. The tool occasionally produced inaccurate logic or non-existent syntax. This taught us a critical lesson: AI outputs cannot be taken at face value and always require rigorous manual verification.

##  3.5 Future Usage
In my future studies and career, I will use GenAI as a learning aid to break down unfamiliar technical concepts and deepening my understanding of newer topics.

##  3.6 Additional Thoughts
GenAI should supplement learning, not replace critical thinking. While everyone should leverage these tools to boost productivity, we must always use our own knowledge and expertise to monitor, evaluate, and check the AI's output.

What do yall think?
