---
title: "Task02"
format: html
editor: visual
---

Task 2: Activation Function Analysis

---

## **2.1 Visualizing `swish()` Activation Function and its Gradient**

```{r}
#| echo: true
library(tidyverse)
```

You can add options to executable code like this

```{r}

#| eval: true
#| echo: true
#| label: fig-swish
#| fig-cap: "Swish Activation Function"
#| fig-width: 7
#| fig-height: 4

swish <- function(x) {
  x / (1 + exp(-x))
}

x <- -10:10
y <- swish(x)
tbl_swish <- tibble(x, y)

ggplot(tbl_swish, aes(x, y)) +
  geom_line(color = "darkblue", linewidth = 1) +
  labs(x = "x", y = "swish(x)")

```

The `echo: false` option disables the printing of code (only output is displayed).

```{r}
#| eval: true
#| echo: true
#| output: true
#| label: fig-d-swish
#| fig-cap: "Gradient of Swish Activation Function"
#| fig-width: 7
#| fig-height: 4

d_swish <- function(x) {
  sigmoid <- 1 / (1 + exp(-x))
  sigmoid + x * sigmoid * (1 - sigmoid)
}

x <- -10:10
y <- d_swish(x)
tbl_d_swish <- tibble(x, y)

ggplot(tbl_d_swish, aes(x, y)) +
  geom_line(color = "darkblue", linewidth = 1) +
  labs(x = "x", y = "d_swish(x)")

```

Unlike ReLU, whose gradient is strictly 0 for negative inputs and 1 for positive inputs, the gradient of Swish is smooth and non-zero for most inputs, with values that gradually transition rather than abruptly cut off. This helps avoid the **dying neuron problem** seen in ReLU, where neurons with negative inputs stop learning entirely. However, Swish is computationally more expensive as it requires calculating the sigmoid function, whereas ReLU involves only a simple threshold operation.

## **2.2 Toy Examples Calling `swish()` and `relu()`**
