---
title: "team13"
format: html
editor: visual
---

## Task 1: EPL

### 1.1 Generating the League Table
```{r}
library(tidyverse)
library(gt)
library(conflicted)

conflicts_prefer(dplyr::filter)
```

```{r}
epl_results <- read_csv(
  "data/epl_results.csv",
  show_col_types = FALSE
)

epl_teams <- read_csv(
  "data/epl_teams.csv",
  show_col_types = FALSE
)
```

```{r}
epl_teams <- epl_teams |>
  select(
    team = 1,
    team_abbr = 2
  ) |>
  mutate(
    team_abbr = str_to_lower(team_abbr)
  )
```

```{r}
matches <- epl_results |>
  pivot_longer(
    cols = -home_team,
    names_to = "away_team",
    values_to = "score"
  ) |>
  filter(score != "-") |>
  separate(
    col = score,
    into = c("home_goals", "away_goals"),
    sep = "-",
    convert = TRUE
  ) |>
  mutate(
    home_team = str_to_lower(home_team),
    away_team = str_to_lower(away_team)
  )

home_performance <- matches |>
  mutate(
    team_abbr = home_team,
    goals_scored = home_goals,
    goals_conceded = away_goals,
    points = case_when(
      home_goals > away_goals ~ 3,
      home_goals == away_goals ~ 1,
      TRUE ~ 0
    )
  ) |>
  select(
    team_abbr,
    goals_scored,
    goals_conceded,
    points
  )

away_performance <- matches |>
  mutate(
    team_abbr = away_team,
    goals_scored = away_goals,
    goals_conceded = home_goals,
    points = case_when(
      away_goals > home_goals ~ 3,
      away_goals == home_goals ~ 1,
      TRUE ~ 0
    )
  ) |>
  select(
    team_abbr,
    goals_scored,
    goals_conceded,
    points
  )

league_table <- bind_rows(
  home_performance,
  away_performance
) |>
  group_by(team_abbr) |>
  summarise(
    points = sum(points),
    goals_scored = sum(goals_scored),
    goals_conceded = sum(goals_conceded),
    .groups = "drop"
  ) |>
  mutate(
    goal_diff = goals_scored - goals_conceded
  ) |>
  left_join(
    epl_teams,
    by = "team_abbr"
  ) |>
  arrange(
    desc(points),
    desc(goal_diff),
    desc(goals_scored)
  ) |>
  mutate(
    rank = row_number()
  ) |>
  select(
    rank,
    team,
    points,
    goal_diff,
    goals_scored
  )

league_table |>
  gt() |>
  tab_header(
    title = md("**English Premier League 2019–2020**"),
    subtitle = "Final Season Results"
  ) |>
  cols_label(
    rank = "Rank",
    team = "Team",
    points = "Points",
    goal_diff = "Goal Difference",
    goals_scored = "Goals Scored"
  ) |>
  cols_align(
    align = "left",
    columns = team
  ) |>
  cols_align(
    align = "center",
    columns = c(rank, points, goal_diff, goals_scored)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = c(rank, team)
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#E7E7E7"),
    locations = cells_body(
      rows = rank %% 2 == 0
    )
  ) |>
  tab_style(
    style = cell_fill(color = "limegreen"),
    locations = cells_body(
      rows = rank <= 4
    )
  ) |>
  tab_style(
    style = cell_fill(color = "red"),
    locations = cells_body(
      rows = rank >= max(rank) - 2
    )
  ) |>
  tab_source_note(
    source_note = md(
      "**Green:** Qualified for Champions League  
       **Red:** Relegated"
    )
  ) |>
  tab_options(
    table.width = pct(100),
    heading.align = "left"
  )
```

### 1.2 Metrics
For the Upset Scoring Index, an upset is defined as scoring more or fewer goals per match against the top six teams than the average achieved by all 14 non-top-six teams. A positive metric value represents above-average scoring against stronger opposition, while a negative value represents below-average scoring.

```{r}
top_6 <- league_table |>
  filter(rank <= 6) |>
  left_join(
    epl_teams,
    by = "team"
  ) |>
  select(
    team_abbr,
    team
  )

bottom_14 <- league_table |>
  filter(rank > 6) |>
  left_join(
    epl_teams,
    by = "team"
  ) |>
  select(
    team_abbr,
    team
  )

top_6_abbr <- top_6$team_abbr
bottom_14_abbr <- bottom_14$team_abbr

top_6_matches <- matches |>
  filter(
    (home_team %in% top_6_abbr &
       away_team %in% bottom_14_abbr) |
      (home_team %in% bottom_14_abbr &
         away_team %in% top_6_abbr)
  )

bottom_14_home <- top_6_matches |>
  filter(home_team %in% bottom_14_abbr) |>
  mutate(
    team_abbr = home_team,
    goals_against_top_6 = home_goals
  ) |>
  select(
    team_abbr,
    goals_against_top_6
  )

bottom_14_away <- top_6_matches |>
  filter(away_team %in% bottom_14_abbr) |>
  mutate(
    team_abbr = away_team,
    goals_against_top_6 = away_goals
  ) |>
  select(
    team_abbr,
    goals_against_top_6
  )

bottom_14_results <- bind_rows(
  bottom_14_home,
  bottom_14_away
)

overall_scoring <- bottom_14_results |>
  summarise(
    total_goals = sum(goals_against_top_6),
    total_matches = n(),
    average_goals = total_goals / total_matches
  )

overall_average <- overall_scoring$average_goals

upset_scoring_table <- bottom_14_results |>
  group_by(team_abbr) |>
  summarise(
    total_goals = sum(goals_against_top_6),
    matches_played = n(),
    average_goals = total_goals / matches_played,
    .groups = "drop"
  ) |>
  mutate(
    metric_value = average_goals - overall_average,
    performance = case_when(
      metric_value > 0 ~ "Above average",
      metric_value < 0 ~ "Below average",
      TRUE ~ "Equal to average"
    )
  ) |>
  left_join(
    epl_teams,
    by = "team_abbr"
  ) |>
  arrange(
    desc(metric_value),
    desc(total_goals)
  ) |>
  mutate(
    metric_rank = row_number()
  ) |>
  select(
    metric_rank,
    team,
    metric_value,
    total_goals,
    matches_played,
    average_goals,
    performance
  )

average_row <- tibble(
  metric_rank = NA_integer_,
  team = "Average",
  metric_value = 0,
  total_goals = mean(upset_scoring_table$total_goals),
  matches_played = mean(upset_scoring_table$matches_played),
  average_goals = overall_average,
  performance = "Average"
)

upset_scoring_table <- rbind(
  upset_scoring_table,
  average_row
) |>
  arrange(
    desc(average_goals)
  )

overall_scoring
```

```{r}
green_fill <- "limegreen"
red_fill <- "red"

dark_green <- adjustcolor(
  green_fill,
  red.f = 0.8,
  green.f = 0.8,
  blue.f = 0.8
)

dark_red <- adjustcolor(
  red_fill,
  red.f = 0.8,
  green.f = 0.8,
  blue.f = 0.8
)

upset_scoring_table |>
  gt() |>
  sub_missing(
    columns = metric_rank,
    missing_text = ""
  ) |>
  tab_header(
    title = md("**Upset Scoring Index**"),
    subtitle = "Goals scored against the EPL Top Six"
  ) |>
  cols_label(
    metric_rank = "Rank",
    team = "Team",
    metric_value = "Metric Value",
    total_goals = "Total Goals",
    matches_played = "Matches",
    average_goals = "Goals per Match",
    performance = "Comparison"
  ) |>
  fmt_number(
    columns = c(
      metric_value,
      total_goals,
      matches_played,
      average_goals
    ),
    decimals = 2
  ) |>
  cols_align(
    align = "left",
    columns = c(team, performance)
  ) |>
  cols_align(
    align = "center",
    columns = c(
      metric_rank,
      metric_value,
      total_goals,
      matches_played,
      average_goals
    )
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = c(metric_rank, team)
    )
  ) |>
  tab_style(
    style = cell_fill(color = green_fill),
    locations = cells_body(
      rows = metric_value > 0
    )
  ) |>
  tab_style(
    style = cell_fill(color = dark_green),
    locations = cells_body(
      rows = metric_value > 0 &
        metric_rank %% 2 == 0
    )
  ) |>
  tab_style(
    style = cell_fill(color = red_fill),
    locations = cells_body(
      rows = metric_value < 0
    )
  ) |>
  tab_style(
    style = cell_fill(color = dark_red),
    locations = cells_body(
      rows = metric_value < 0 &
        metric_rank %% 2 != 0
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#CFE8FF"),
    locations = cells_body(
      rows = metric_value == 0 &
        performance == "Average"
    )
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      rows = metric_value == 0 &
        performance == "Average"
    )
  ) |>
  tab_source_note(
    source_note = md(
      "**Positive:** Above the overall scoring average  
       **Negative:** Below the overall scoring average  
       **Blue:** Overall average"
    )
  ) |>
  tab_options(
    table.width = pct(100),
    heading.align = "left"
  )
```

## Task 2: Investment Market Price

## Task 3: Visualising Singapore Tourist Spending

## Task 4: Reflection on GenAI use
